---
title: "DATA205 Homework 7"
author: "Emma King"
date: "3/13/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(odbc)
library(DBI)
library(RSQLite)
```


## A. Create a SQL database called olympics.db that contains one table for each .csv.

```{r}
athlete_events <- readr::read_csv('https://raw.githubusercontent.com/rgriff23/Olympic_history/refs/heads/master/data/athlete_events.csv')

host_city_locations <- readr::read_csv('https://raw.githubusercontent.com/rgriff23/Olympic_history/refs/heads/master/data/host_city_locations.csv')
```

```{r}
olympics <- dbConnect(RSQLite::SQLite(), ":memory:")
copy_to(olympics, athlete_events)
copy_to(olympics, host_city_locations)
```



## B. Write queries to answer the following questions:

### 1. How many records are in athlete_events?

There are 271,116 records in the athlete_events table.

```{sql connection=olympics}
SELECT COUNT(*) AS num_records
FROM athlete_events

```


### 2. How many different athletes competed in Olympics held in London?

16,941 athletes have competed in Olympics hosted by London.

```{sql connection=olympics}
SELECT COUNT(DISTINCT ID) AS num_athletes
FROM athlete_events
WHERE City = 'London'

```


### 3. Which Olympics host city is furthest north?

Lillehammer is the northernmost Olympics host city.

```{sql connection=olympics}
SELECT City
FROM host_city_locations
ORDER BY lat DESC
LIMIT 1

```


### 4. What is the list of host cities (with no duplicates)?

The following is a list of unique Olympics host cities.

```{sql connection=olympics}
SELECT DISTINCT City
FROM host_city_locations

```


```{r}
distinct_host_cities <- dbGetQuery(olympics, '
SELECT DISTINCT City
FROM host_city_locations
') 

write.csv(distinct_host_cities, "C:/Users/jaria/Downloads/distinct_host_cities.csv")
```


```{r}
dbDisconnect(olympics)
```